---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true


---
<style type="text/css">
p.import { color:green;}
p.import1{color:red;}
</style>


## Loading and preprocessing the data

```{r load}
load<-read.csv("activity/activity.csv", header=TRUE,stringsAsFactors=FALSE)
data<-load

```


## What is mean total number of steps taken per day?

```{r mean}
aggmean<-aggregate(as.numeric(data$steps),list(data$date),mean,na.action=na.omit)
aggsum<-aggregate(as.numeric(data$steps),list(data$date),sum,na.rm=TRUE)

names(aggsum)<-c("Date","Total Steps")
names(aggmean)<-c("Date","Average Steps")

meanperdaymean<-mean(aggmean$"Average Steps", na.rm=TRUE)

meanperalldays<-mean(aggsum$"Total Steps",na.rm=TRUE)
total<-sum(aggsum$"Total Steps",na.rm=TRUE)
 medsteps<-median(aggsum$"Total Steps", na.rm=TRUE)
```
### <p class=import>Total Steps taken in the 2 months: `r total` steps</p>

### <p class=import>There are an average of `r meanperdaymean` steps per interval. </p>

### <p class=import>There are an average of `r meanperalldays` steps taken per day. </p>

### <p class=import>There is a median of `r medsteps` steps taken per day. </p>

### Histogram of total steps.
```{r total steps histogram}
hist(aggsum$"Total Steps",breaks=length(aggsum$"Date"))
```



## What is the average daily activity pattern?


Take the average of each interval.
Plot a time series of averages for 1-288 intervals.
```{r average steps per interval}
aggin<-aggregate(as.numeric(data$steps),list(data$interval),mean, na.rm=TRUE)
names(aggin)<-c("Interval","Average Steps")
 times<-ts(aggin$"Average Steps",start=1,end=288, frequency=1)
plot(times)

```

What interval has the greatest average?
```{r interval with largest average}
mx<-max(aggin$"Average Steps")
mxs<-aggin[aggin$"Average Steps"==mx,]
```
### <p class=import>The largest average for any interval is `r mx` average steps taken interval per day. </p>
### <p class=import>The interval with the greatest average steps per day is interval `r mxs$Interval` at `r mxs$"Average Steps"` steps taken per day. </p>








## Imputing missing values


```{r missing values}
comp<-complete.cases(data)
incomplete<-data[!comp,]
inc<-length(incomplete[[1]])
```
### <p class=import>The dataset has `r inc` observations with missing values.</p>


```{r fill in missing values}

findint<-function(int,indf){

	steps<-0
	for(n in 1:length(indf[[1]]) )
	{ 
		if(indf[n,"Interval"]==int){
		steps<-indf[n,"Average Steps"]
	   	}
	}
	steps
}

fill<-function(df,indf){

	filled<-data.frame()
	for(i in 1:length(df[[1]]))
	{
		if(is.na(df$steps[i]))
		{
			df$steps[i]<- findint(df[i,"interval"], indf)
		}
	filled<-rbind(filled,df[i,])
	}
	filled
}
d<-data
i<-aggin
completedcases<-fill(d,i)
```
## Strategy:<p class=import1>Fill in missing values with the average number of steps for each missing interval.</p>

### Histogram of total steps in complete cases dataset
```{r histogram of complete cases}
completedaggsum<-aggregate(as.numeric(completedcases$steps),list(completedcases$date),sum,na.rm=TRUE)
names(completedaggsum)<-c("Date","Total Steps")
hist(completedaggsum$"Total Steps",breaks=length(completedaggsum$"Date"))

```
```{r mean and median steps with completed cases}
completedmeanperalldays<-mean(completedaggsum$"Total Steps",na.rm=TRUE)
completedtotal<-sum(completedaggsum$"Total Steps",na.rm=TRUE)
 completedmedsteps<-median(completedaggsum$"Total Steps", na.rm=TRUE)
```
### <p class=import>Total Steps taken in the 2 months: `r completedtotal` steps in the filled in dataset.</p>

### <p class=import>There are an average of `r completedmeanperalldays` steps taken per day in the filled in data set. </p>

### <p class=import>There is a median of `r completedmedsteps` steps taken per dayin the filled in dataset. </p>

### Do these values differ from the estimates from the first part of the assignment? 
The mean and median of the filled in dataset are close to the mean and median of the original data.
### What is the impact of imputing missing data on the estimates of the total daily number of steps?
By replacing missing values with the average steps taken for each interval, the mean is now equal to the median. Adding the averages into the missing values decreased the skew of the distribution of steps.



## Are there differences in activity patterns between weekdays and weekends?

Split add Weekend/Weekday factor to data set.
```{r factor the data set by weekend/weekday}

getday<-function(obj){weekdays(as.Date(obj))}

days<-sapply(completedcases$date,getday)
casesdays<-cbind(completedcases,days)

week<-c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday")
weekend<-c("Sunday","Saturday")

getweekend<-function(d, ke){
	result<-character()
	for(i in 1:length(ke)){
		if(d==ke[i])
			result<-"Weekend"
		else
			result<-"Weekday"
	}
	result
}

weekcat<-sapply(casesdays$days,getweekend, weekend)
weekcases<-cbind(casesdays,weekcat)
str(weekcases)
weekframes<-split(weekcases,weekcases$weekcat)
```

Get average steps per interval on weekdays
```{r weekday steps}
weekdayaggin<-aggregate(as.numeric(weekframes[[1]]$steps),list(weekframes[[1]]$interval),mean, na.rm=TRUE)

names(weekdayaggin)<-c("Interval","Average Steps")
weekdaytimes<-ts(weekdayaggin$"Average Steps",start=1,end=288, frequency=1)
weekdayaggin$weekcat<-"Weekday"
```

Get average steps per interval on weekends
```{r weekend steps}
weekendaggin<-aggregate(as.numeric(weekframes[[2]]$steps),list(weekframes[[2]]$interval),mean, na.rm=TRUE)

names(weekendaggin)<-c("Interval","Average Steps")
weekendtimes<-ts(weekendaggin$"Average Steps",start=1,end=288, frequency=1)
weekendaggin$weekcat<-"Weekend"
```

### Plot averages for weekend and weekdays
```{r}
weekaggin<-rbind(weekdayaggin,weekendaggin)
library(lattice)
xyplot(weekaggin$"Average Steps" ~ weekaggin$Interval | weekaggin$weekcat,type="l", main="Average Steps taken for 288 intervals",layout=c(1,2),xlab="Interval",ylab="Avg Steps taken")

```

